---
title: "Vern_phenotyping_2019"
output: html_document
---

# Load the packages:
```{r, message=F}
library(readxl)
library(dplyr)

# for the date data in the excel file
library(janitor)

#
library(plyr)
library(reshape2)

#
library(ggplot2)

#
library(emmeans)
```


# Load the data:
```{r, warning=F}
pheno_16_8 <- read_excel("C:/Users/Po-Kai/Box Sync/Runcie_lab/Vernalization_project/Vernalization_project_PKH_part/Phenotyping_201904/Phenotyping.xlsx", 
    sheet = "16-8", col_types = c("text", "text", "numeric", "numeric", "numeric", "numeric", "skip", "skip", "skip"), na = "NA")

pheno_12_12 <- read_excel("C:/Users/Po-Kai/Box Sync/Runcie_lab/Vernalization_project/Vernalization_project_PKH_part/Phenotyping_201904/Phenotyping.xlsx", 
    sheet = "12-12", col_types = c("text", "text", "numeric", "numeric", "numeric", "numeric", "skip", "skip", "skip"), na = "NA")


#Dealing with the date data in the excel file  
pheno_16_8$Bolting_date <- excel_numeric_to_date(pheno_16_8$Bolting_date)  
pheno_16_8$Flowering_date <- excel_numeric_to_date(pheno_16_8$Flowering_date)  

pheno_12_12$Bolting_date <- excel_numeric_to_date(pheno_12_12$Bolting_date)  
pheno_12_12$Flowering_date <- excel_numeric_to_date(pheno_12_12$Flowering_date) 


# Calculate days to bolting and flowering
pheno_16_8$Days_to_bolting = as.numeric(pheno_16_8$Bolting_date) - as.numeric(as.Date('2019-07-9'))
pheno_16_8$Days_to_flowering = as.numeric(pheno_16_8$Flowering_date) - as.numeric(as.Date('2019-07-9'))
pheno_12_12$Days_to_bolting = as.numeric(pheno_12_12$Bolting_date) - as.numeric(as.Date('2019-07-9'))
pheno_12_12$Days_to_flowering = as.numeric(pheno_12_12$Flowering_date) - as.numeric(as.Date('2019-07-9'))

#
pheno_16_8$Trt <- factor(pheno_16_8$Trt, levels = c("1d", "7d", "22d", "35d", "NA"))
pheno_12_12$Trt <- factor(pheno_12_12$Trt, levels = c("1d", "7d", "22d", "35d", "NA"))
```


# For leaf number: 
```{r}
# (16/8 day/night cycle)
leaf_16_8 <- select(pheno_16_8, Genotype, Trt, ends_with("number"))
leaf_16_8$Total_leaves <- leaf_16_8$Rosette_number + leaf_16_8$Cauline_number

# Different ways to view the data
## log2: the differences between the bars = the ratio
#leaf_16_8[,3:5] = log2(leaf_16_8[,3:5])
## Reciprocal: The rate (think: 70 leaves to flower = 1 leaf: 1/70 flower)
#leaf_16_8[,3:5] = 1/(leaf_16_8[,3:5])

# re-organize (melt) the data and use ddply to get the mean, SD, and SEM 
m_leaf_16_8 <- melt(leaf_16_8, id.vars=c("Genotype", "Trt"))
summary_leaf_16_8 <- ddply(m_leaf_16_8, c("Genotype", "Trt", "variable"), summarise,
      mean = mean(value, na.rm=TRUE), sd = sd(value, na.rm=TRUE),
      sem = sd(value, na.rm=TRUE)/sqrt(length(value)))

# Standard deviation of the mean as error bar: Rosette leaf
p <- ggplot(summary_leaf_16_8 %>% filter(variable=="Rosette_number"), aes(x=Trt, y=mean, fill=Genotype)) + 
   geom_bar(stat="identity", position=position_dodge()) +
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.2,
                 position=position_dodge(.9))
  
p + scale_fill_brewer(palette="Paired") + theme_minimal() + xlab("Treatment") + ylab("Rosette Leaf number") + ggtitle("Rosette Leaf Number under 16/8 h day/night cycles")

# Standard deviation of the mean as error bar: Cauline leaf
p <- ggplot(summary_leaf_16_8 %>% filter(variable=="Cauline_number"), aes(x=Trt, y=mean, fill=Genotype)) + 
   geom_bar(stat="identity", position=position_dodge()) +
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.2,
                 position=position_dodge(.9))
  
p + scale_fill_brewer(palette="Paired") + theme_minimal() + theme_minimal() + xlab("Treatment") + ylab("Cauline Leaf number") + ggtitle("Cauline Leaf Number under 16/8 h day/night cycles")

# Standard deviation of the mean as error bar: Total leaves
p <- ggplot(summary_leaf_16_8 %>% filter(variable=="Total_leaves"), aes(x=Trt, y=mean, fill=Genotype)) + 
   geom_bar(stat="identity", position=position_dodge()) +
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.2,
                 position=position_dodge(.9))
  
p + scale_fill_brewer(palette="Set2") + theme_minimal() + theme_minimal() + xlab("Treatment") + ylab("Total Leaf number") + ggtitle("Total Leaf Number under 16/8 h day/night cycles")


######################################
# (12/12 day/night cycle)
leaf_12_12 <- select(pheno_12_12, Genotype, Trt, ends_with("number"))
leaf_12_12$Total_leaves <- leaf_12_12$Rosette_number + leaf_12_12$Cauline_number

# Different ways to view the data
#leaf_12_12[,3:5] = log2(leaf_12_12[,3:5])
#leaf_12_12[,3:4] = 1/(leaf_12_12[,3:4])

# re-organize (melt) the data and use ddply to get the mean, SD, and SEM 
m_leaf_12_12 <- melt(leaf_12_12, id.vars=c("Genotype", "Trt"))
summary_leaf_12_12 <- ddply(m_leaf_12_12, c("Genotype", "Trt", "variable"), summarise,
      mean = mean(value, na.rm=TRUE), sd = sd(value, na.rm=TRUE),
      sem = sd(value, na.rm=TRUE)/sqrt(length(value)))

# Standard deviation of the mean as error bar: Rosette leaf
p <- ggplot(summary_leaf_12_12 %>% filter(variable=="Rosette_number"), aes(x=Trt, y=mean, fill=Genotype)) + 
   geom_bar(stat="identity", position=position_dodge()) +
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.2,
                 position=position_dodge(.9))
p + scale_fill_brewer(palette="Paired") + theme_minimal() + theme_minimal() + xlab("Treatment") + ylab("Rosette Leaf number") + ggtitle("Rosette Leaf Number under 12/12 h day/night cycles")

# Standard deviation of the mean as error bar: Cauline leaf
p <- ggplot(summary_leaf_12_12 %>% filter(variable=="Cauline_number"), aes(x=Trt, y=mean, fill=Genotype)) + 
   geom_bar(stat="identity", position=position_dodge()) +
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.2,
                 position=position_dodge(.9))

p + scale_fill_brewer(palette="Paired") + theme_minimal() + xlab("Treatment") + ylab("Cauline Leaf number") + ggtitle("Cauline Leaf Number under 12/12 h day/night cycles")

# Standard deviation of the mean as error bar: Total leaves
p <- ggplot(summary_leaf_12_12 %>% filter(variable=="Total_leaves"), aes(x=Trt, y=mean, fill=Genotype)) + 
   geom_bar(stat="identity", position=position_dodge()) +
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.2,
                 position=position_dodge(.9))
  
p + scale_fill_brewer(palette="Set2") + theme_minimal() + xlab("Treatment") + ylab("Total Leaf number") + ggtitle("Total Leaf Number under 12/12 h day/night cycles")

```
(1) Use the leaf number:
- The number: 1d < 7d < 22d < 35d
- The difference between Col and Col-FRI: 1d ~ 7d < 22d < 35d (The later vernalization, the bigger difference between Col and Col-FRI.)


(2) Use the 1/leaf number:
- The number: 1d > 7d > 22d > 35d
- The difference between Col and Col-FRI: Similar between groups. (The difference of flowering rate remain the same under different treatments.)


(3) Use the log2(leaf):
- The number: 1d < 7d < 22d < 35d
- The difference between Col and Col-FRI: 1d ~ 7d ~ 22d < 35d. (The difference means Col/Col-FRI.)
(But what is the biological meaning??? I think it can represent how many times the leaf number of Col does Col-FRI need to flower.)


# More analyses on leaf number
```{r}
# The differences between Col and Col-FRI:
m1 = lm(mean~Trt + variable + Genotype:variable:Trt,summary_leaf_16_8)
# You can compare the results (the mean) of emmeans to the summary tables I made above.
#emmeans(m1,specs = 'Genotype',by=c('Trt','variable'))
# This will return a list, so use the as.data.frame to combine the list.
#pairs(emmeans(m1,specs = 'Genotype',by=c('Trt','variable')))
as.data.frame(pairs(emmeans(m1,specs = 'Genotype',by=c('Trt','variable'))))


# For 12/12 h cycle:
as.data.frame(pairs(emmeans(lm(mean~Genotype:Trt:variable,summary_leaf_12_12),specs = 'Genotype',by=c('Trt','variable'))))


```
For Total leaves, the difference between Col and Col-FRI:
(1) 16/8 h > 12/12 h: 1d, 7d, 22d
(2) 16/8 h < 12/12 h: 35d 

Thus, the effects of LD (16/8) is smaller in 35d trt but bigger in all others. 
> Does that support Dan's model?


# For days to flower:
```{r}
# (16/8 day/night cycle)
days_16_8 <- select(pheno_16_8, Genotype, Trt, starts_with("Days"))


# Different ways to view the data
## log2: the differences between the bars = the ratio
#days_16_8[,3:4] = log2(days_16_8[,3:4])
## Reciprocal: The rate (think: 70 leaves to flower = 1 leaf: 1/70 flower)
#days_16_8[,3:4] = 1/(days_16_8[,3:4])

# re-organize (melt) the data and use ddply to get the mean, SD, and SEM 
m_days_16_8 <- melt(days_16_8, id.vars=c("Genotype", "Trt"))
summary_days_16_8 <- ddply(m_days_16_8, c("Genotype", "Trt", "variable"), summarise,
      mean = mean(value, na.rm=TRUE), sd = sd(value, na.rm=TRUE),
      sem = sd(value, na.rm=TRUE)/sqrt(length(value)))

# Standard deviation of the mean as error bar: Days_to_bolting
p <- ggplot(summary_days_16_8 %>% filter(variable=="Days_to_bolting"), aes(x=Trt, y=mean, fill=Genotype)) + 
   geom_bar(stat="identity", position=position_dodge()) +
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.2,
                 position=position_dodge(.9))
  
p + scale_fill_brewer(palette="Paired") + theme_minimal() + xlab("Treatment") + ylab("Days") + ggtitle("Days to bolting under 16/8 h day/night cycles")

# Standard deviation of the mean as error bar: Days_to_flowering
p <- ggplot(summary_days_16_8 %>% filter(variable=="Days_to_flowering"), aes(x=Trt, y=mean, fill=Genotype)) + 
   geom_bar(stat="identity", position=position_dodge()) +
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.2,
                 position=position_dodge(.9))
  
p + scale_fill_brewer(palette="Paired") + theme_minimal() + theme_minimal() + xlab("Treatment") + ylab("Days") + ggtitle("Days to flowering under 16/8 h day/night cycles")


######################################
# (12/12 day/night cycle)
days_12_12 <- select(pheno_12_12, Genotype, Trt, starts_with("Days"))


# Different ways to view the data
## log2: the differences between the bars = the ratio
#days_12_12[,3:4] = log2(days_12_12[,3:4])
## Reciprocal: The rate (think: 70 leaves to flower = 1 leaf: 1/70 flower)
#days_12_12[,3:4] = 1/(days_12_12[,3:4])

# re-organize (melt) the data and use ddply to get the mean, SD, and SEM 
m_days_12_12 <- melt(days_12_12, id.vars=c("Genotype", "Trt"))
summary_days_12_12 <- ddply(m_days_12_12, c("Genotype", "Trt", "variable"), summarise,
      mean = mean(value, na.rm=TRUE), sd = sd(value, na.rm=TRUE),
      sem = sd(value, na.rm=TRUE)/sqrt(length(value)))

# Standard deviation of the mean as error bar: Days_to_bolting
p <- ggplot(summary_days_12_12 %>% filter(variable=="Days_to_bolting"), aes(x=Trt, y=mean, fill=Genotype)) + 
   geom_bar(stat="identity", position=position_dodge()) +
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.2,
                 position=position_dodge(.9))
  
p + scale_fill_brewer(palette="Paired") + theme_minimal() + xlab("Treatment") + ylab("Days") + ggtitle("Days to bolting under 12/12 h day/night cycles")

# Standard deviation of the mean as error bar: Days_to_flowering
p <- ggplot(summary_days_12_12 %>% filter(variable=="Days_to_flowering"), aes(x=Trt, y=mean, fill=Genotype)) + 
   geom_bar(stat="identity", position=position_dodge()) +
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.2,
                 position=position_dodge(.9))
  
p + scale_fill_brewer(palette="Paired") + theme_minimal() + theme_minimal() + xlab("Treatment") + ylab("Days") + ggtitle("Days to flowering under 12/12 h day/night cycles")


```


# Exclude the NV trt
```{r}
# Standard deviation of the mean as error bar: Days_to_bolting
p <- ggplot(summary_days_16_8 %>% filter(variable=="Days_to_bolting")%>% subset(Trt != "NA"), aes(x=Trt, y=mean, fill=Genotype)) + 
   geom_bar(stat="identity", position=position_dodge()) +
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.2,
                 position=position_dodge(.9))
  
p + scale_fill_brewer(palette="Paired") + theme_minimal() + xlab("Treatment") + ylab("Days") + ggtitle("Days to bolting under 16/8 h day/night cycles")

# Standard deviation of the mean as error bar: Days_to_flowering
p <- ggplot(summary_days_16_8 %>% filter(variable=="Days_to_flowering") %>% subset(Trt != "NA"), aes(x=Trt, y=mean, fill=Genotype)) + 
   geom_bar(stat="identity", position=position_dodge()) +
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.2,
                 position=position_dodge(.9))
  
p + scale_fill_brewer(palette="Paired") + theme_minimal() + theme_minimal() + xlab("Treatment") + ylab("Days") + ggtitle("Days to flowering under 16/8 h day/night cycles")

##################################################################
# Standard deviation of the mean as error bar: Days_to_bolting
p <- ggplot(summary_days_12_12 %>% filter(variable=="Days_to_bolting") %>% subset(Trt != "NA"), aes(x=Trt, y=mean, fill=Genotype)) + 
   geom_bar(stat="identity", position=position_dodge()) +
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.2,
                 position=position_dodge(.9))
  
p + scale_fill_brewer(palette="Paired") + theme_minimal() + xlab("Treatment") + ylab("Days") + ggtitle("Days to bolting under 12/12 h day/night cycles")

# Standard deviation of the mean as error bar: Days_to_flowering
p <- ggplot(summary_days_12_12 %>% filter(variable=="Days_to_flowering") %>% subset(Trt != "NA"), aes(x=Trt, y=mean, fill=Genotype)) + 
   geom_bar(stat="identity", position=position_dodge()) +
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.2,
                 position=position_dodge(.9))
  
p + scale_fill_brewer(palette="Paired") + theme_minimal() + theme_minimal() + xlab("Treatment") + ylab("Days") + ggtitle("Days to flowering under 12/12 h day/night cycles")

```


# Comparing the 12/12-h and 16/8-h conditions:
```{r}
ggplot(summary_leaf_12_12,aes(x=Trt,y=mean)) + geom_point(color='red') + geom_point(data = summary_leaf_16_8,color='black') + facet_grid(Genotype~variable)

# Different code, the same result:
#ggplot(summary_leaf_12_12,aes(x=Trt,y=mean)) + geom_point(color='red') + facet_wrap(Genotype~variable) + geom_point(data=summary_leaf_16_8,color='blue')

ggplot(summary_days_12_12,aes(x=Trt,y=mean)) + geom_point(color='red') + geom_point(data = summary_days_16_8,color='black') + facet_grid(Genotype~variable)


```

(1) The NV condition: except for the cauline leaf number, other parameters show the delay of flowering under 12/12-h condition.

(2) The rosette leaf number: Generally, the number of leaves are slightly  higher under 12/12-h than under 16/8-h for both Col and Col-FRI. (But the differences are not very big) 

(3) The total leaf number: Generally, the number of leaves are slightly  higher under 12/12-h than under 16/8-h for both Col and Col-FRI. (But the differences are not very big) 

(4) However, both the rosette and total leaf number are identical under 35d using Col.

(5) The obvious difference between 16/8-h and 12/12-h conditions: Col, 1d; Col-FRI, 1d; Col-FRI, 35d.

(6) The flowering time seems to show more differences: most of the plants under 12/12-h flower later than those under 16/8-h. Photosynthesis?

(7) Interestingly, the NV Col can flower as fast as 1d Col under 16/8-h condition, but even slower than 7d under 12/12-h condition. That means the LD can complement the vern requirement even for Col genotype. Also, even Col is considered as the vernalization-not-required ecotype, it seems the vern can accerate the flowering of Col under 12/12-d condition. Think: the FLC-independent pathway!

