---
title: "Vern_201908_7d_Trt_qPCR"
output: html_document
---
Include the data generated in 20190819 and 20190831


# Load the packages
```{r, message=FALSE}
library(readxl)
library(dplyr)
library(plyr)
library(reshape2)
library(ggplot2)
```


# Load the data from excel
```{r}
Raw_data <- read_excel("C:/Users/Po-Kai/Box Sync/Runcie_lab/Vernalization_project/Vernalization_project_PKH_part/qPCR_data/2019_project/201908.xlsx", 
    sheet = "Raw_data")

Raw_data_1212vern <- read_excel("C:/Users/Po-Kai/Box Sync/Runcie_lab/Vernalization_project/Vernalization_project_PKH_part/qPCR_data/2019_project/20190930_1212vern.xlsx", 
    sheet = "Raw_data")


#
#UBC <- read_excel("C:/Users/Po-Kai/Box Sync/Runcie_lab/Vernalization project/Vernalization_project_PKH_part/qPCR_data/2019_project/20190904.xlsx", sheet = "UBC")

#FT <- read_excel("C:/Users/Po-Kai/Box Sync/Runcie_lab/Vernalization project/Vernalization_project_PKH_part/qPCR_data/2019_project/20190904.xlsx", sheet = "FT")


```


# dCT and fold change
```{r}
#summary_Raw <- ddply(Raw_data, c("Sample_Name","Target_Name"), summarise, mean = mean(CT, na.rm=TRUE))

summary_Raw <- ddply(Raw_data, c("Sample_Name","Target_Name"), summarise,
      mean = mean(CT, na.rm=TRUE), sd = sd(CT, na.rm=TRUE))


FT_UBC <- dcast(summary_Raw, formula = Sample_Name ~ Target_Name, value.var = "mean")
FT_UBC$dCT <- FT_UBC$FT - FT_UBC$UBC
FT_UBC$fold_change <- 2^(-(FT_UBC$dCT))

# Add the leaf rank
FT_UBC$leaf_rank <- factor(c("65", "60", "55", "50","45","40","35","30","25","65", "60", "55", "50","45","40","35","30","25","20","70","65", "60", "55", "50","45","40","35","30","25","20"), level=c("20","25","30","35","40","45","50","55","60","65","70"))
```


# Biological replicate
```{r}
dCT_final <- ddply(FT_UBC, c("leaf_rank"), summarise,
      mean = mean(fold_change, na.rm=TRUE), sd = sd(fold_change, na.rm=TRUE))

#ddply(FT_UBC %>% select(c("Sample_Name","dCT","leaf_rank")), c("leaf_rank"), summarise,
#      mean = mean(dCT, na.rm=TRUE), sd = sd(dCT, na.rm=TRUE))


p <- ggplot(dCT_final, aes(x=leaf_rank, y=mean, fill=leaf_rank)) + 
   geom_bar(stat="identity", position=position_dodge()) +
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.2,
                 position=position_dodge(.9))

p + scale_fill_brewer(palette="Paired") + theme_minimal() + labs(x = "Leaf rank", y = "Relative expression to UBC")


```
However, the bar plot may not be suitable for presenting the data. 


```{r}
# Check where is the boundary of each individual plant:
FT_UBC$leaf_rankN = as.numeric(as.character(FT_UBC$leaf_rank))
which(diff(FT_UBC$leaf_rankN)> 0)
FT_UBC$Sample = c(rep(1,9),rep(2,10),rep(3,11))


# Trial and error:
#ggplot(FT_UBC,aes(x=leaf_rankN,y=fold_change)) + geom_point(aes(color = Sample)) + geom_smooth(aes(color=Sample,group = Sample),method='loess',span=.3)
#ggplot(FT_UBC,aes(x=leaf_rankN,y=fold_change)) + geom_point(aes(color = Sample)) + geom_smooth(aes(color=Sample,group = Sample),method='loess',span=.5)


# Convert the "Sample" to factor:
FT_UBC$Sample = factor(FT_UBC$Sample)
ggplot(FT_UBC,aes(x=leaf_rankN,y=fold_change)) + geom_point(aes(color = Sample)) + geom_smooth(aes(color=Sample,group = Sample),method='loess',span=.5)

ggplot(FT_UBC,aes(x=leaf_rankN,y=fold_change)) + geom_point(aes(color = Sample)) + geom_smooth(aes(color=Sample,group = Sample),method='loess')


# Can we pool the data together? Something like:
ggplot(FT_UBC,aes(x=leaf_rankN,y=fold_change)) + geom_point(aes(color = Sample)) + geom_smooth(method='loess',span=.5)
```
(1) How to choose the curve?
-> 20190925, Dan: Use the last one

(2) Why there is CI even for the each biological replicate?
-> Dan: loess uses several points to generate the curve, so the CI is from the points it uses.


20190930: Add the reslts of 12h/12h chamber, 35d
```{r}
# dCT and fold change
summary_Raw_1212 <- ddply(Raw_data_1212vern, c("Sample_Name","Target_Name"), summarise,
      mean = mean(CT, na.rm=TRUE), sd = sd(CT, na.rm=TRUE))


FT_UBC_1212 <- dcast(summary_Raw_1212, formula = Sample_Name ~ Target_Name, value.var = "mean")
FT_UBC_1212$dCT <- FT_UBC_1212$FT - FT_UBC_1212$UBC
FT_UBC_1212$fold_change <- 2^(-(FT_UBC_1212$dCT))

# Add the leaf rank
FT_UBC_1212$leaf_rank <- factor(c("35", "30", "25", "20", seq(from=65, to=20, by=-5), "18", seq(from=70, to=20, by=-5), seq(from=70, to=40, by=-5)))

FT_UBC_1212$leaf_rankN = as.numeric(as.character(FT_UBC_1212$leaf_rank))
FT_UBC_1212$Sample = c(rep(3,4),rep(1,11),rep(2,11), rep(3,7))
FT_UBC_1212$Sample = factor(FT_UBC_1212$Sample)

# Plots:
ggplot(FT_UBC_1212,aes(x=leaf_rankN,y=fold_change)) + geom_point(aes(color = Sample)) + geom_smooth(aes(color=Sample,group = Sample),method='loess',span=.5)

ggplot(FT_UBC_1212,aes(x=leaf_rankN,y=fold_change)) + geom_point(aes(color = Sample)) + geom_smooth(aes(color=Sample,group = Sample),method='loess')


# Can we pool the data together? Something like:
ggplot(FT_UBC_1212,aes(x=leaf_rankN,y=fold_change)) + geom_point(aes(color = Sample)) + geom_smooth(method='loess',span=.5)


```
(1) Anyway to merge two curve together?
